# Prediction Assignment Writeup - Quantifying the Weight Lifting Exercise Activity
`r opts_chunk$set(cache=TRUE)`
`r opts_chunk$set(warning=FALSE)`
## Synopsis

Human Activity Recognition - HAR has emerged as a key research activity in the recent years, but the emphasis was more on "how much" than "how well" a certain activity was done. 

Using data from the Weight Lifting Exercise Dataset of the HAR project [1] we apply machine learning techniques to predict with a high accuracy the correctness of the exercise execution.

## Data Processing

The data is read in data.frame structures as it can be seen below:

```{r loadingData}
pmlte<-read.table("pml-testing.csv", dec=".", fill=TRUE, quote="\"", 
                 row.names=NULL,
                 header=TRUE, sep=",")
pmltr<-read.table("pml-training.csv", dec=".", fill=TRUE, quote="\"", 
                  row.names=NULL,
                  header=TRUE, sep=",")
```

An exploratory approach of the data is taken, in order to gain more insight into the judicious choice of the features to be considered.

If we calculate the variance of all the features (the columns), we can see that many of them are "NA" or close to 0. We choose to ignore them and build a new dataset with a limited number of features - those that 'matter'/can potentially influence the prediction outcome.

```{r filteringSignificatColumns}
apply(pmltr,2,var)
data<-subset(pmltr, select=c("roll_belt", "pitch_belt","yaw_belt","total_accel_belt",
"accel_belt_x","accel_belt_y","accel_belt_z", "magnet_belt_x","magnet_belt_y",
"magnet_belt_z","roll_arm","pitch_arm","yaw_arm","total_accel_arm","gyros_arm_x",
"accel_arm_x","accel_arm_y","accel_arm_z","magnet_arm_x","magnet_arm_y","magnet_arm_z",
"roll_dumbbell","pitch_dumbbell","yaw_dumbbell","total_accel_dumbbell","gyros_dumbbell_x",
"gyros_dumbbell_z","accel_dumbbell_x","accel_dumbbell_y","accel_dumbbell_z",
"magnet_dumbbell_x","magnet_dumbbell_y","magnet_dumbbell_z","roll_forearm","pitch_forearm",
"yaw_forearm","total_accel_forearm","gyros_forearm_y","gyros_forearm_z","accel_forearm_x",
"accel_forearm_y","accel_forearm_z","magnet_forearm_x","magnet_forearm_y","magnet_forearm_z",
"classe"))
```

The total computation time for the training part will be decreased by this limitation of the number of the features.

We create 'training' and 'testing' datasets with this data.

```{r createDataPartitions}
library(caret)
intrain<-createDataPartition(y=data$classe, p=0.75,list=FALSE)
training<-data[intrain,]
testing<-data[-intrain,]
```

We choose a "random forest" machine learning algorithm because we have a classification problem with many of parameters and this choice could have a good accuracy from the first try.

The 'testing' dataset will be used for cross-validation, after building the model.

For performance reasons, the 'randomForest' library was used, as it can be seen below.

```{r prediction}
library("randomForest")
library("doParallel")
registerDoParallel(cores=detectCores())
modrf<-randomForest(classe ~ ., data = training)
prf<-predict(modrf,newdata=testing)
```

## Results
We obtain an overall accuracy of 99.5%. The "random forest" approach gives a very good result for this problem.

```{r results}
confusionMatrix(prf, testing$classe)
```

[1] Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. *Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements*. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6.
